---
title: "Project2"
format: html
editor: visual
---

## Find Data You Are Interested In

```{r}
# Loading appropriate Packages
library(tidyverse)
library(ggplot2)
```

For my project I chose to use the Melbourne Housing Market full data set downloaded from Kaggle

```{r}
# Reading in Melbourne Houseing data (Full dataset)
data <- read_csv("Melbourne_housing_FULL.csv")
head(data)
```

## Prepare for Your App

The data set appears to have missing values. We will omit lines that contain the missing elements. Additionally, we'll change the units for Type. 

```{r}
# Cleaning data set 
data <- na.omit(data)
head(data)
```

```{r}
# Changing the units for Type
data <- data |>
  mutate(Type = recode(Type,
                       "h" = "House",
                       "u" = "Unit",
                       "t" = "Townhouse"))
unique(data$Type)
```


### Produce numerical and graphical summaries to investigate the data.

1) One-way contingency tables

```{r}
# One-way contingency table for Suburb
data |>
  group_by(Suburb) |>
  summarize(counts = n())
```

```{r}
# One-way contingency table for SellerG (seller's name)
data |>
  group_by(SellerG) |>
  summarize(counts = n())
```

2) Two-way contingency tables

```{r}
# Two-way contingency table for SellerG and Suburb
# Shows us how many homes each seller sold per suburb
data |>
  group_by(SellerG, Suburb) |>
  summarize(counts = n()) 
```

```{r}
# Two-way contingency table for SellerG and Type
# Shows us what type of homes sellers prefer
data |>
  group_by(SellerG, Type) |>
  summarize(counts = n()) 
```

3) Numerical summaries (means, medians, sds, etc.) for quantitative variables at levels of categorical
variables

```{r}
# Numerical summary of Price, Rooms, and yearBuilt by Type 

data |>
  group_by(Type) |>
  select(Price, Rooms, YearBuilt) |>
  summarize(across(where(is.numeric), 
                   list("mean" = ~ mean(.x), "median" = ~ median(.x, na.rm = TRUE), "sd" = ~ sd(.x), "IQR" = ~ IQR(.x)), 
                   .names = "{.fn}_{.col}"))
```

```{r}
# Numerical summary of Price, Rooms, and Car by Suburb 
# Shows us details about home/garage size and their price per suburb
data |>
  group_by(Suburb) |>
  select(Price, Rooms, Car) |> 
  summarize(across(where(is.numeric), 
                   list("mean" = ~ mean(.x), "median" = ~ median(.x, na.rm = TRUE), "sd" = ~ sd(.x), "IQR" = ~ IQR(.x)), 
                   .names = "{.fn}_{.col}"))
```

4) Create at least six plots to visualize the data

```{r}
# Scatter Plot with faceting: Price and Rooms across Type
sp <- ggplot(data, aes(x = Rooms, y = Price, color = Type))
sp + geom_point() +
  labs(x = "Rooms", y = "Price", 
       title = "Scatter Plot: Home Price and Number of Rooms by Type") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ Type)
```

```{r}
# Loading appropriate packages
library(ggridges)
library(viridis)

# ggridges: Home Price by Council Area
ggplot(data, aes(x = Price, y = CouncilArea)) +
  geom_density_ridges_gradient(aes(fill = ..x..), scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Price", option = "C") +
  labs(title = "Home Price by Council Area", x = "Home Price", y = "City Council") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
bc <- ggplot(data, aes(x = Regionname, fill = Type))
bc + geom_bar(position = "stack") +
  labs(title = "Stacked Bar Chart of Home Type by Region", 
       x = "Region", 
       y = "Count", 
       fill = "Home Type") +
  theme(plot.title = element_text(hjust = 0.5)) + # Center title
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate labels
```

```{r}
# 2D bin Plot for type of home sold per seller in postal code 3067
data2 <- data[data$Postcode == 3067, ] # Too much data
h <- ggplot(data2, aes(x = SellerG, y = Type))
h + geom_bin2d(binwidth = c(0.25, 500), alpha = 0.5, color = "black", fill = "blue", linetype = "solid", linewidth = 0.5) +
  labs(
    title = "Type of Home Sold per Seller in Postal Code 3067",
    x = "Seller",
    y = "Home Type") +
  theme(plot.title = element_text(hjust = 0.5))  # Center the title
```

```{r}
h <- ggplot(data2, aes(x = Price, fill = SellerG))
h + geom_histogram(binwidth = 50000, color = "black", alpha = 0.7) +  # Adjust `binwidth` as needed
  labs(
    x = "Price",
    y = "Count",
    title = "Prices of Homes Sold Across Sellers in Postal Code 3067") +
  scale_fill_discrete(name = "Sellers") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)) +# Rotates x-axis labels for readability
  facet_wrap(~ SellerG)


```


```{r}
ggplot(data, aes(x = Lattitude, y = Longtitude)) +
  geom_bin2d(bins = 50) +
  scale_fill_viridis_c() +
  labs(title = "2D Density Plot of Coordinates",
       x = "X Coordinate",
       y = "Y Coordinate") +
  theme(plot.title = element_text(hjust = 0.5))
```

